Can we distinguish modes of selective interactions using linkage disequilibrium?

### 2021-03-14 Manuscript.Rmd
edits to methods and figures

Many well-known properties of two-locus dynamics and equilibrium LD come from
early work on the multi-locus diffusion approximation [@Kimura1955-qe;
@Hill1968-vu; @Ohta1969-ie; @Ohta1971-yd]. This includes the result that
genome-wide averages of signed LD are expected to be zero under neutrality.
Under a two-locus biallelic model, where the left locus allows alleles $A$ and
$a$ and the right locus has alleles $B$ and $b$, the standard covariance
measure of LD is defined as $D = f_{AB} - f_{A}f_{B}$ where $f_{AB}$ is the
haplotype frequency of double-derived types carrying both $A$ and $B$, and
$f_{A}$ and $f_{B}$ are the marginal frequencies of the derived alleles at each
locus. This covariance decays due to both drift and recombination at a rate
proportional to the inverse of the effective population size and the distance
separating loci: \[\E[D]_{t+1} = \left(1 - \frac{1}{2N_e(t)} - r
\right)\E[D]_t.\] While $\E[D] = 0$, the variance of $D$ is non-zero, and
@Ohta1971-yd presented their groundbreaking result that the variance of $D$
under neutrality and steady-state demography, normalized by the joint
heterozygosity of the two loci, is \begin{equation} (\#eq:ohta) \sigma_d^2 =
\frac{\E[D^2]}{\E[f_A(1-f_A)f_B(1-f_B)]}\approx\frac{5 + \rho / 2}{11 + 13\rho
/ 2 + \rho^2 / 2}, \end{equation} where $\rho = 4N_e r$.

@@

To include selection, however, there have been relatively few advances beyond
the Monte Carlo simulation approach taken by @Hill1966-gv, albeit we now have
access to somewhat more powerful computational resources and more sophisticated
software for performing flexible forward simulation (e.g., @Thornton2014-pn,
@Haller2019-vm). Analytic results for two-locus distributions under selection
are notoriously difficult, with a few notable flashes of progress. For example,
@McVean2007-xd considered the effect of a recent sweep on patterns of LD in
loci nearby or spanning the locus under selection, and in an impressive recent
paper, @Good2020-wj presented analytic solutions for patterns of LD between
rare mutations under additive selection with epistasis. Nonetheless, such
approaches are typically confined to steady-state demography and constrained
selection models.

@@

Under neutrality, a number of approaches exist to compute $\Psi_n$, including
the recursion due to @Golding1984-pu and @Ethier1990-tb, or more recent
numerical approaches to the two-locus coalescent [@Kamm2016-ag] or diffusion
approximation [@Ragsdale2017-gg]. Practically, selection can only be included
using the forward-in-time diffusion equation [@Kimura1955-qe; @Hill1966-gv],
where a standard approach is to first solve for the continuous distribution
$\psi$ of the density of two-locus haplotype configurations in the full
population, and then integrate $\psi$ against the multinomial sampling function
to obtain $\Psi_n$.

@@

The system of linear ordinary differential equations for the entries of
$\Psi_n$takes the form
\begin{equation}
(\#eq:system)
{\Psi}_n^{t+1}(i, j, k; t) = 
\mathcal{D}_{N(t)}\Psi_n^t
+ \mathcal{R}_{r}\Psi_{n+1}^t
+ \mathcal{U}_{u}\Psi_n^t
+ \mathcal{S}_{s_A, s_B, \ldots, h_A, h_B, \ldots}\Psi_{n+2}^t.
\end{equation}
Here, $\mathcal{D}_{N(t)}$ is a sparse linear operator accounting for drift
with population size $N(t)$, $\mathcal{R}$ accounts for recombination
with per-generation recombination probability $r$ between the two loci,
$\mathcal{U}$ accounts for mutation, either under an infinite sites or
biallelic reversible mutation model, and $\mathcal{S}$ accounts for
selection.

The moment system for $\Psi_n$ can be derived directly from the diffusion
approximation, or it can be found through a more intuitive process of tracking
the dynamics of allelic states of a sample of size of $n$ from the full
population. We assume $n \ll N_e$, and $r$ and $s$ are $O(1/Ne)$ so that
multiple coalescence, recombination, or selective events are rare in any given
generation. For example, for drift events, we account for the probability that
two lineages within the $n$ tracked lineages at time $t+1$ coalesced in the
previous generation and, given a coalescence event, the probability that
haplotype counts changed due to this coalescence event [see Supporting
Information; @Jouganous2017-pq; @Ragsdale2019-nt]. In typical diffusion
approximation fashion, we multiple through by $2N_{ref}$ so that time
is measured in $2N_{ref}$ generations, and we consider scaled parameters
$\rho = 4Nr$, $\theta = 4Nu$, and $\gamma=2Ns$.

@@

In the absence of selection and for fully linked loci (i.e., $\rho=0$), the
system is closed and can be solved exactly. However, for nonzero recombination
or selection, the entries of $\Psi_n$ rely on the slightly larger frequency
spectra with sample sizes $n+1$ (for recombination and additive selection) or
$n+2$ (for non-additive selection). This is because if a recombination event
occurs within one of $n$ lineages being tracked by $\Psi_n$, we need to draw an
additional lineage from the full population to recombine with that chosen
lineage, thus requiring $\Psi_{n+1}^t$ to find $\Psi_n^{t+1}$. Selection events
similarly require extra lineages from the full population, which replace a
chosen lineage that fails to reproduce with probability proportional to its
selection coefficient, assuming negative selection.

This requirement of extra lineages for nonzero recombination and selection
means that the system in \@ref(eq:system) is not closed, so that we need a
moment-closure approximation to solve for $\Psi_n$. As in @Ragsdale2019-nt, a
jackknife approximation is used to estimate $\Psi_{n+l}$, for $l=1$ or $2$,
from $\Psi_n$ [@Jouganous2017-pq], so that $\hat{\Psi}_{n+l}(i, j, k) =
\mathcal{J}_{n, l}\Psi_n$, although other accurate closure approximations are
possible [@Friedlander2021-bs]. This emits a closed approximate system,
\begin{equation}
(\#eq:closed-system)
\dot{\hat\Psi}_n(i, j, k; t) = 
\mathcal{D}_{\nu(t)}\hat\Psi_n(t)
+ \mathcal{R}_{\rho}\mathcal{J}_{n, 1}\hat\Psi_{n}(t)
+ \mathcal{U}_{\theta}\hat\Psi_n(t)
+ \mathcal{S}_{\gamma, h}\mathcal{J}_{n, 2}\hat\Psi_{n}(t).
\end{equation}

The jackknife approximation, which approximates an entry $\Psi_{n+l}(i,j,k)$
using nearby entries in $\Psi_n$, is more accurate for larger sample sizes,
creating a tension between efficiency and accuracy: larger sample sizes result
in more accurate solutions, as error in the jackknife is diminished, but
computational complexity also grows rapidly in the number of entries of
$\Psi_n$, which is $O(n^3)$. Accuracy and runtime are shown in Figure
\@ref(fig:jackknife). In the results presented in this paper, sample sizes
between $n=30$ and $n=80$ are used.

@@

To include selection, we consider a model where we draw lineages uniformly from
the previous generation, but keep lineages with probability proportional to
their fitness. In the absence of dominance, selection reduces to a haploid
model, with acceptance and rejection probabilities depending on the fitnesses
of each haploid copy, where haplotype $Ab$ has fitness $1 + s_{A}$, $aB$ has
fitness $1 + s_{B}$, and $AB$ has fitness $1 + s_{AB}$. We assume the doubly
ancestral haplotype $ab$ has fitness $1$, so fitnesses are relative to that of
$ab$ haplotypes. The standard multiplicative fitness function assumes that
$s_{AB} \approx s_{A} + s_{B}$, and a general epistasis function can be written
as $s_{AB} = (s_{A} + s_{B}) (1 + \epsilon)$, so that $\epsilon > 0$ implies
synergistic epistasis, while $\epsilon < 0$ implies antagonistic epistasis.

To obtain the recursion equation under selection we consider drawing $n$
lineages from generation $t$, which has has an expected sampling distribution
of haplotype counts given by $\Psi_n^t$. However, assuming $s\leq0$ for each
haplotype, each of those sampled lineages has probability of being rejected
equal to the relative fitness of its haplotype state. If a lineage is rejected,
a replacement is drawn from the full population. Under the assumption that
$ns_{\cdot\cdot} \ll 1$, the probability that more than one selection event
occurs in any given generation is negligibly small, so that the case of
multiple simultaneous rejections can be ignored. Then $Psi_n^{t+1}$ relies only
on $\Psi_n^t$ and $\Psi_{n+1}^t$ for additive selection. The full selection
operator $\mathcal{S}$ for additive selection is given in the Supporting
Information.

To account for dominance within loci, or other general forms of two-locus
selection, the selection operator no longer reduces to individual haplotypes,
but instead we need to know the state of two-locus genotypes. For example, the
fitness of an individual carrying an $Ab$ haplotype depends on whether their
second haplotype is $ab$, $Ab$, $aB$, or $AB$. We can therefore assign a
selection coefficient to each possible diploid configuration, $s_{AB/AB}$,
$s_{AB/Ab}$, and so on. Assuming that the doubly homozygous ancestral $ab/ab$
genootype has relative fitness 1, this gives nine possible unique selection
coefficients in the most general two-locus selection model. Note that $AB/ab$
and $Ab/aB$ genotypes need not have the same selection coefficient, which
allows for simulation under a gene-based model of dominance (Table
\@ref(tab:selmodels)).

@@

Using the annotated variant call format (VCF) files from the phase 3
@1000_Genomes_Project_Consortium2015-zq (Thousand Genomes) data release, I
subset the genotype VCFs of the same release to autosomal variants that were
annotated as either synonymous or nonsynonymous (including both missense and
more damaging mutations such as stop gains, stop losses, and so on). I further
subset to samples within each non-admixed population in the African, European,
and East Asian continental groups. Signed LD is sensitive to ancestral state
identification, so I only kept sites for which ancestral alleles were estimated
with high confidence in both the VCF info fields and the Thousand Genomes human
ancestor reconstructed from a phylogeny of six primates.

@@

# Acknowledgements
- Simon Gravel
- Kevin Thornton
- Alex Diaz-Papkovich
- Mashaal Sohail
- Matthias Steinrücken

@@

(ref:fig1cap) \textbf{The two-locus sampling distribution and additive
selection.} (A and B) The distribution of $AB$ haplotype counts in a sample
size of 30, in which we observe 10 $A$ alleles at the left locus, and 10 $B$
alleles at the right locus. (C) The decay of normalized squared LD
($\sigma_d^2$) with recombination distance ($\rho=4N_er$) for pairs of variants
with a given selection coefficient. The dashed line is the neutral expectation
from @Ohta1971-yd, given in Equation \@ref(eq:ohta). (D) The decay of
$\sigma_d^1=\E[D]/ \E[p(1-p)q(1-q)]$ for pairs of variants. The
negative LD effect of Hill-Robertson interference is strongest for intermediate
selection, with $\gamma\sim-1$ to $-5$.

@@

(ref:fig2cap) \textbf{Additive selection with epistasis.} (A, B) Synergistic
epistasis decreases $D^2$, while antagonistic epistasis increases $D^2$. (C, D)
Epistasis expected signed LD away from zero, an effect which can extend over
long recombination distances. Here, $\gamma$ is the selection coefficient for
derived alleles $A$ and $B$, assuming selection is equal at the two loci.
$\epsilon=0$ is the case of pure additive selection with no epistasis, and
$\gamma_{AB}=(\gamma_A + \gamma_B)(1+\epsilon)$. Dashed lines are expectations
under neutrality for both $\sigma_d^2$ and $\sigma_d^1$.

@@

(ref:fig4cap) **The effects of demography on signed LD.** Left column:
simulations under two toy models, one with an instantaneous expansion in the
past, and the other with a 10-fold reduction in size followed by recovery.
Right column: simulations under population size histories inferred using Relate
[@Speidel2019-nj] applied to populations from
@1000_Genomes_Project_Consortium2015-zq. In all simulations, $\gamma$ is set to
$-2$ at both loci. (A-B) Model illustrations. Note that the differences in
scale. (C-D) Additive selection with no epistasis, showing fully linked
($\rho=0$) and more loosely linked ($\rho=5$) pairs. Bottleneck events result
in larger negative values of signed LD, but population size changes are not
expected to change the sign of mean LD. (E-F) Synergistic ($\epsilon > 0$) and
antagonistic ($\epsilon < 0$) epistasis. Again, bottleneck events result in
larger signed LD. Here, $\epsilon=1/2$ is a model with $\gamma_A=-2$,
$\gamma_B=-2$ and $\gamma_{AB}=-6$, and $\epsilon=-1/2$ is a model with
$\gamma_A=-2$, $\gamma_B=-2$ and $\gamma_{AB}=-2$. (G-H) Dominance within loci
results in negative LD, while dominance acting at the gene level results in
positive LD. Here, $h=0.1$. Again, bottlenecks result in larger average LD in
both cases.

@@

(ref:fig5cap) **LD within and between protein-coding genes.** (A-B) Pairs of
variants in coding regions partitioned by classes of mutations.
Loss-of-function mutations are defined as stop and start gains and losses,
frameshifts, and other nonsense mutations. In all populations, we observe
positive mean signed LD, although the confidence interval of each measurement
overlaps with zero. Similarly, we observe mean negative LD for pairs of
loss-of-function variants within genes. There are only $\sim200-300$ pairs of
loss of function variants that fall within the same gene across the populations
tested here, causing large uncertainty in each measurement. Gray lines indicate
2 standard errors of the measurement, via bootstrap. (C-F) While pairs of
synonymous and missense mutations within genes are consistently positive, pairs
between genes are not consistent in their sign across populations. LD for all
populations analyzed here are shown in Figure \@ref(fig:genespops).

@@

(ref:fig6cap) **LD within and between annotated domains in protein-coding
genes.** Considering only variants that fall within annotated domains in
protein-coding genes, pairs of mutations that fall within the same gene display
larger positive LD than pairs that lie in different domains. This pattern is
most striking for missense mutations, for which pairs within domains have
average signed LD that is consistently larger than for pairs that span two
domains, or for pairs of synonymous mutations that are within the same domain.
The slightly positive LD that we observe for missense and synonymous mutations
gene-wide (Figure \@ref(fig:fig4)A) is largely driven by the subset of pairs
mutations that fall within the same annotated domain. LD for all populations
analyzed here are shown in Figure \@ref(fig:domainspops).

>>>>

Many well-known properties of two-locus dynamics and equilibrium LD come from
early work on the multi-locus diffusion approximation [@Kimura1955-qe;
@Hill1968-vu; @Ohta1969-ie; @Ohta1971-yd]. This includes the result that
genome-wide averages of signed LD are expected to be zero under neutrality.
Under a two-locus biallelic model, where the left locus allows alleles $A$ and
$a$ and the right locus allows alleles $B$ and $b$, the standard covariance
measure of LD is defined as $D = f_{AB} - f_{A}f_{B}$, where $f_{AB}$ is the
haplotype frequency of double-derived types carrying both $A$ and $B$, and
$f_{A}$ and $f_{B}$ are the marginal frequencies of the derived alleles at each
locus. This covariance decays due to both drift and recombination at a rate
proportional to the inverse of the effective population size and the distance
separating loci [@Hill1968-vu]: \[\E[D]_{t+1} = \left(1 - \frac{1}{2N_e(t)} - r
\right)\E[D]_t.\] While $\E[D] = 0$, the variance of $D$ is non-zero, and
@Ohta1971-yd presented their groundbreaking result that the variance of $D$
under neutrality and steady-state demography, normalized by the joint
heterozygosity of the two loci, is \begin{equation} (\#eq:ohta) \sigma_d^2 =
\frac{\E[D^2]}{\E[f_A(1-f_A)f_B(1-f_B)]}\approx\frac{5 + \rho / 2}{11 + 13\rho
/ 2 + \rho^2 / 2}, \end{equation} where $\rho = 4N_e r$.

@@

To include selection, however, there have been relatively few advances beyond
the Monte Carlo simulation approach taken by @Hill1966-gv, albeit now with more
powerful computational resources and sophisticated software for performing
flexible forward simulation (e.g., @Thornton2014-pn, @Haller2019-vm). Analytic
results for two-locus distributions under selection are notoriously difficult,
with a few notable flashes of progress. For example, @McVean2007-xd considered
the effect of a recent sweep on patterns of LD between neutral loci near the
locus under selection, and in a recent paper, @Good2020-wj presented analytic
solutions for patterns of LD between rare mutations under additive selection
with epistasis. Nonetheless, such approaches are typically confined to
steady-state demography and constrained selection models.

@@

Under neutrality, a number of approaches exist to compute $\Psi_n$, including
the recursion due to @Golding1984-pu and @Ethier1990-tb, or more recent
numerical approaches to the two-locus coalescent [@Kamm2016-ag] or diffusion
approximation [@Ragsdale2017-gg]. Selection is most easily included using the
forward-in-time diffusion equation [@Kimura1955-qe; @Hill1966-gv], where a
standard approach is to first solve for the continuous distribution $\psi$ of
the density of two-locus haplotype configurations in the full population, and
then integrate $\psi$ against the multinomial sampling function to obtain
$\Psi_n$.

@@

The system of linear ordinary differential equations for the entries of
$\Psi_n$ takes the form
\begin{equation}
(\#eq:system)
{\Psi}_n^{t+1}(i, j, k; t) = 
\mathcal{D}_{N(t)}\Psi_n^t
+ \mathcal{R}_{r}\Psi_{n+1}^t
+ \mathcal{U}_{u}\Psi_n^t
+ \mathcal{S}_{s_A, s_B, \ldots, h_A, h_B, \ldots}\Psi_{n+2}^t.
\end{equation}
Here, $\mathcal{D}_{N(t)}$ is a sparse linear operator accounting for drift
with population size $N(t)$, $\mathcal{R}$ accounts for recombination
with per-generation recombination probability $r$ between the two loci,
$\mathcal{U}$ accounts for mutation, either under an infinite sites or
biallelic reversible mutation model, and $\mathcal{S}$ accounts for
selection.

The moment system for $\Psi_n$ can be derived directly from the diffusion
approximation, or it can be found through a more intuitive process of tracking
the dynamics of allelic states of a sample of size of $n$ from the full
population. We assume $n \ll N_e$, and $r$ and $s$ are $O(1/Ne)$ so that
multiple coalescence, recombination, or selective events within the $n$
lineages are rare in any given generation. For example, for drift events, we
account for the probability that two lineages within the $n$ tracked lineages
at time $t+1$ coalesced in the previous generation and, given a coalescence
event, the probability that haplotype counts changed due to this coalescence
event [see Supporting Information; @Jouganous2017-pq; @Ragsdale2019-nt]. In
typical diffusion approximation fashion, we multiple through by $2N_{ref}$ so
that time is measured in $2N_{ref}$ generations, and we consider scaled
parameters $\rho = 4Nr$, $\theta = 4Nu$, and $\gamma=2Ns$.

@@

In the absence of selection and for fully linked loci (i.e., $\rho=0$), the
system is closed and can be solved exactly. However, for nonzero recombination
or selection, the entries of $\Psi_n$ rely on the slightly sampling
distributions with sample sizes $n+1$ (for recombination and additive
selection) or $n+2$ (for non-additive selection). This is because if a
recombination event occurs within one of $n$ lineages being tracked by
$\Psi_n$, we need to draw an additional lineage from the full population to
recombine with that chosen lineage, thus requiring $\Psi_{n+1}^t$ to find
$\Psi_n^{t+1}$. Selection events similarly require extra lineages from the full
population, which replace a chosen lineage that fails to reproduce with
probability proportional to its selection coefficient, assuming negative
selection.

This requirement of extra lineages for nonzero recombination and selection
means that the system in \@ref(eq:system) is not closed, so that we need a
moment-closure approximation to solve for $\Psi_n$. As in @Ragsdale2019-nt, a
jackknife approximation is used to estimate $\Psi_{n+l}$, for $l=1$ or $2$,
from $\Psi_n$ [following the single-locus closure introduced in
@Jouganous2017-pq], so that $\hat{\Psi}_{n+l}(i, j, k) = \mathcal{J}_{n,
l}\Psi_n$, although other accurate closure approximations are possible
[@Friedlander2021-bs]. This emits a closed approximate system,
\begin{equation}
(\#eq:closed-system)
\dot{\hat\Psi}_n(i, j, k; t) = 
\mathcal{D}_{\nu(t)}\hat\Psi_n(t)
+ \mathcal{R}_{\rho}\mathcal{J}_{n, 1}\hat\Psi_{n}(t)
+ \mathcal{U}_{\theta}\hat\Psi_n(t)
+ \mathcal{S}_{\gamma, h}\mathcal{J}_{n, 2}\hat\Psi_{n}(t).
\end{equation}

The jackknife approximation, which approximates an entry $\Psi_{n+l}(i,j,k)$
using nearby entries in $\Psi_n$, is more accurate for larger sample sizes
(Figure \@ref(fig:jackknife)), creating a tension between efficiency and
accuracy: larger sample sizes result in more accurate solutions, as error in
the jackknife is diminished, but computational complexity also grows rapidly in
the number of entries of $\Psi_n$, which is $O(n^3)$. Accuracy and runtime are
shown in Figure \@ref(fig:jackknife). In the results presented in this paper,
sample sizes between $n=30$ and $n=80$ are used.

@@

To include selection, we consider a model where we draw lineages uniformly from
the previous generation, but keep lineages with probability proportional to
their fitness. In the absence of dominance, selection reduces to a haploid
model, with acceptance and rejection probabilities depending on the fitnesses
of each haploid copy, where haplotype $Ab$ has fitness $1 + s_{A}$, $aB$ has
fitness $1 + s_{B}$, and $AB$ has fitness $1 + s_{AB}$. We assume the doubly
ancestral haplotype $ab$ has fitness $1$, so fitnesses are relative to that of
$ab$ haplotypes. The standard multiplicative fitness function assumes that
$s_{AB} \approx s_{A} + s_{B}$, and the effect of epistasis can be written as
$s_{AB} = (s_{A} + s_{B}) (1 + \epsilon)$, so that $\epsilon > 0$ implies
synergistic epistasis, while $\epsilon < 0$ implies antagonistic epistasis.

To obtain the recursion equation under selection we consider drawing $n$
lineages from generation $t$, which has has an expected sampling distribution
of haplotype counts given by $\Psi_n^t$. However, assuming $s\leq0$ for each
derived haplotype, each of those sampled lineages has probability of being
rejected equal to the relative fitness of its haplotype state. If a lineage is
rejected, a replacement is drawn from the full population. Under the assumption
that $ns_{\cdot\cdot} \ll 1$, the probability that more than one selection
event occurs in any given generation is negligibly small, so that the case of
multiple simultaneous rejections can be ignored. Then $Psi_n^{t+1}$ relies only
on $\Psi_n^t$ and $\Psi_{n+1}^t$ for additive selection. The full selection
operator $\mathcal{S}$ for additive selection is given in the Supporting
Information.

To account for dominance within loci, or other general forms of two-locus
selection, the selection operator no longer reduces to individual haplotypes,
but instead we need to know the state of two-locus genotypes. For example, the
fitness of an individual carrying an $Ab$ haplotype depends on whether their
second haplotype is $ab$, $Ab$, $aB$, or $AB$. We can therefore assign a
selection coefficient to each possible diploid configuration, $s_{AB/AB}$,
$s_{AB/Ab}$, and so on. Assuming that the doubly homozygous ancestral $ab/ab$
genotype has relative fitness 1, this gives nine possible unique selection
coefficients in the most general two-locus selection model. Note that $AB/ab$
and $Ab/aB$ genotypes need not have the same selection coefficient, which
allows for simulation under a gene-based model of dominance (Table
\@ref(tab:selmodels)).

@@

Using the annotated variant call format (VCF) files from the phase 3
@1000_Genomes_Project_Consortium2015-zq (Thousand Genomes) data release, I
subset the genotype VCFs of the same release to autosomal variants that were
annotated as either synonymous or nonsynonymous (including both missense and
more damaging mutations such as stop gains, stop losses, and so on). I further
subset to samples within each non-admixed population in the African, European,
and East Asian continental groups. Signed LD is sensitive to ancestral state
identification, so I only kept sites for which ancestral alleles were estimated
with high confidence in both the VCF info field and the Thousand Genomes human
ancestor reconstructed from a phylogeny of six primates.

@@

# Acknowledgements
- Discussions: Simon Gravel, Alex Diaz-Papkovich, Matthias Steinrücken, Eric Friedlander
- Comments on ms:

Asked from comments from
- Kevin Thornton
- Mashaal Sohail
- Simon Gravel
- Matthias Steinrücken

@@

(ref:fig1cap) \textbf{The two-locus sampling distribution and additive
selection.} (A and B) The distribution of $AB$ haplotype counts in a sample
size of 30, in which we observe 10 $A$ alleles at the left locus, and 10 $B$
alleles at the right locus. (C) The decay of normalized squared LD
($\sigma_d^2$) with recombination distance ($\rho=4N_er$) for pairs of variants
with a given selection coefficient. The dashed line is the neutral expectation
from @Ohta1971-yd, given in Equation \@ref(eq:ohta). (D) The decay of
$\sigma_d^1=\E[D]/ \E[p(1-p)q(1-q)]$ for pairs of variants. The negative LD
effect of Hill-Robertson interference is strongest for intermediate selection,
with $\gamma\sim-1$ to $-5$, where $\gamma=4N_es$.

@@

(ref:fig2cap) \textbf{Additive selection with epistasis.} (A, B) Synergistic
epistasis decreases $D^2$, while antagonistic epistasis increases $D^2$. (C, D)
Epistasis expected signed LD away from zero, an effect which can extend over
large recombination distances. Here, $\gamma$ is the selection coefficient for
derived alleles $A$ and $B$, assuming selection is equal at the two loci.
$\epsilon=0$ is the case of pure additive selection with no epistasis, and
$\gamma_{AB}=(\gamma_A + \gamma_B)(1+\epsilon)$. Dashed lines are expectations
under neutrality for both $\sigma_d^2$ and $\sigma_d^1$.

@@

(ref:fig4cap) **The effects of demography on signed LD.** Left column:
simulations under two toy models, one with an instantaneous expansion in the
past, and the other with a 10-fold reduction in size followed by recovery.
Right column: simulations under population size histories inferred using Relate
[@Speidel2019-nj] applied to populations from
@1000_Genomes_Project_Consortium2015-zq. In all simulations, $\gamma$ is set to
$-2$ at both loci. (A-B) Model illustrations. Note the differences in
scale. (C-D) Additive selection with no epistasis, showing fully linked
($\rho=0$) and more loosely linked ($\rho=5$) pairs. Bottleneck events result
in larger negative values of signed LD, but population size changes are not
expected to change the sign of mean LD. (E-F) Synergistic ($\epsilon > 0$) and
antagonistic ($\epsilon < 0$) epistasis. Again, bottleneck events result in
larger signed LD. Here, $\epsilon=1/2$ is a model with $\gamma_A=-2$,
$\gamma_B=-2$ and $\gamma_{AB}=-6$, and $\epsilon=-1/2$ is a model with
$\gamma_A=-2$, $\gamma_B=-2$ and $\gamma_{AB}=-2$. (G-H) Dominance within loci
results in negative LD, while dominance acting at the gene level results in
positive LD. Here, $h=0.1$. Again, bottlenecks result in larger average LD in
both cases.

@@

(ref:fig5cap) **LD within and between protein-coding genes.** (A-B) Pairs of
variants in coding regions partitioned by classes of mutations.
Loss-of-function mutations are defined as stop and start gains and losses and
other nonsense mutations of high impact. In all populations, we observe
positive mean signed LD for missense and synonymous variants, although the
confidence interval of each measurement overlaps with zero. Similarly, we
observe mean negative LD for pairs of loss-of-function variants within genes.
There are only $\sim200-300$ pairs of loss-of-function variants that fall
within the same gene across the populations tested here, causing large
uncertainty in each measurement. Gray lines indicate 2 standard errors of the
measurement, computed by bootstrapping of sets of genes. (C-F) While pairs of
synonymous and missense mutations within genes are consistently positive, pairs
between genes are not consistent in their sign across populations. LD for all
populations analyzed here are shown in Figure \@ref(fig:genespops).

@@

(ref:fig6cap) **LD within and between annotated domains in protein-coding
genes.** (A) Considering only variants that fall within annotated domains in
protein-coding genes, pairs of mutations that fall within the same gene display
larger positive LD than pairs that lie in different domains. This pattern is
most striking for missense mutations, for which pairs within domains have
average signed LD that is consistently larger than for pairs that span two
domains, or for pairs of synonymous mutations that are within the same domain.
(B-D) The slightly positive LD that we observe for missense and synonymous
mutations gene-wide (Figure \@ref(fig:fig4)A) is largely driven by the subset
of pairs mutations that fall within the same annotated domain. LD for all
populations analyzed here are shown in Figure \@ref(fig:domainspops).
